Improving content discovery through combining linked data and data mining techniques

### 2015-11-30 report/report.tex
Big progress on 3rd chapter.

\section{RDF and Feature Extraction}

@@

@@

@@

@@

@@

        IRI -> RDF [label="extract"];
        RDF -> Features [label="generate"];
      }
    \end{dot2tex}
  \end{center}
  \caption{Null feature generator \label{fig:gen-null}}
\end{figure}

@@

There are two clear axes along which we can improve this pipeline: expand
the size of the RDF graph we are extracting and improve how we convert this
graph into features. In the first instance, we can consider the former and add
a single pipeline stage for expanding the RDF graph.

@@

        IRI -> RDFa [label="fetch"];
        RDFa -> RDF [label="parse"];
        RDF -> Features [label="generate"];
      }
    \end{dot2tex}
  \end{center}
  \caption{Semantic web content miner\label{fig:gen-rdfa}}
\end{figure}

Figure~\ref{fig:gen-rdfa}

\begin{sidewaysfigure}[h]
  \begin{center}
    \begin{dot2tex}[dot,options=-t math,autosize,pgf]
      digraph g {
        rankdir=LR;

@@

        IRI -> RDFa [label="fetch"];
        RDFa -> RDF [label="parse"];
        RDF -> RDF1 [label="DBPedia"];
        RDF -> RDF2 [label="DBPedia"];
        RDF -> RDFp [label="\cup"];
        RDF1 -> RDFp [label="\cup"];
        RDF2 -> RDFp [label="\cup"];
        RDFp -> Features [label="generate"];
      }
    \end{dot2tex}
  \end{center}
  \caption{Semantic web content miner with DBPedia enrichment}
\end{sidewaysfigure}

>>>>

\section{Data Mining}

TODO

\section{RDF and Feature Extraction}
\label{sec:rdf-and-features}

@@

TODO

@@

\subsection{Definitions}

@@

\subsection{Identity Graph}

@@

\begin{figure}[h]
  \begin{center}
    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.7]
      digraph g {
        rankdir=LR;

@@

        IRI -> RDF [label="extract"];
        RDF -> Features [label="(IRI)"];
      }
    \end{dot2tex}
  \end{center}
  \caption{Null feature generator \label{fig:gen-null}}
\end{figure}

@@

There are three clear axes along which we can improve this pipeline:
\emph{extract} more RDF data knowing only an item's IRI,
expand or \emph{enrich} an existing RDF graph
and then improve how we \emph{generate} features for data mining.
In the first instance, we can consider the former and add
a single pipeline stage for expanding the RDF graph.

\subsection{RDF Extraction}
\label{sec:rdf-extraction}

@@

        IRI -> RDFa [label="derefernce"];
        RDFa -> RDF [label="parse"];
        RDF -> Features [label="(IRI)"];
      }
    \end{dot2tex}
  \end{center}
  \caption{Semantic web data extraction\label{fig:gen-rdfa}}
\end{figure}

Figure~\ref{fig:gen-rdfa} depicts the pipeline with a simple dereference
step added. Note that the feature set generated is still the singleton
relation with only the IRI value. Thus the next step should be to add a step
that improves the feature set generation.

\subsection{Feature Set Generation}

Paulheim and F\"urnkranz\cite{paulheim2012unsupervised} described a number of
SPARQL queries for generating feature sets from RDF data, which could inspire
a simple query such as that shown in Listings~\ref{lst:simple-sparql}. This
query generates a boolean \texttt{true} value for any properties that match
and implies \texttt{false} for those that do not.

\begin{lstlisting}[label=lst:simple-sparql,caption={Generates field \texttt{content\_?p\_?v} with value \texttt{true}},language=sparql]
SELECT ?p ?v
WHERE { ?iri ?p ?v . }
\end{lstlisting}

As also noted by Paulheim
and F\"urnkranz, this overlooks the
\emph{open world assumption}\cite{russel2010artificial}. However, application
of clustering algorithms on binary data can employ asymmetric distance metrics
such as Jaccard similarity coeffcient\cite{witten2005data}, which notably
avoids deriving similarity from negative values. That is, two content items
lacking a particular property will contribute no information about their
(dis)similarity. Thus we safely avoid inadvertently grouping together one item
that genuinely lacks the property with another that indeed has the property,
but we lack the positive assertion thereof in the data extracted.

\begin{figure}[h]
  \begin{center}
    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.7]
      digraph g {
        rankdir=LR;

        IRI -> RDFa [label="derefernce"];
        RDFa -> RDF [label="parse"];
        RDF -> Features [label="content\_?p\_?v"];
      }
    \end{dot2tex}
  \end{center}
  \caption{Semantic web content extraction with basic SPARQL feature generation\label{fig:gen-rdfa-basic}}
\end{figure}

The basic pipeline in Figure~\ref{fig:gen-rdfa} can thus be augmented with this
basic feature extraction to produce the pipeline depicated in
Figure~\ref{fig:gen-rdfa-basic}.

\subsection{RDF Enrichment}

The third and final direction in which this data pipeline can be improved is
in terms of data enrichment. A simple strategy here is to repeat the
deferencing used in Section~\ref{sec:rdf-extraction}, but for each IRI
found as the object of a triple in which the initial IRI is the subject.
Formally:

$$
g \in G \models \exists p, o . (IRI, p, o) \in g \rightarrow g' = deref(o)
$$

That RDF graphs can be modelled as mathematical sets as in
Section~\ref{sec:rdf-and-features} means we can express a graph enriched this
way as a \emph{union} of the initial graph with each graph returned from all
dereferencing:

$$
g \in G \models g' = g \cup \bigcup \: \{deref(o) \: | \: (IRI, p, o) \in g\}
$$

Figure~\ref{fig:object-deref}

\begin{sidewaysfigure}[h]
  \begin{center}
    \begin{dot2tex}[dot,options=-t math,autosize,pgf,scale=0.7]
      digraph g {
        rankdir=LR;

@@

        IRI -> RDFa [label="dereference"];
        RDFa -> RDF [label="parse"];
        RDF -> RDF1 [label="derefence/parse"];
        RDF -> RDF2 [label="dereference/parse"];
        RDF -> RDFp [label="\cup"];
        RDF1 -> RDFp [label="\cup"];
        RDF2 -> RDFp [label="\cup"];
        RDFp -> Features [label="content\_?p\_?v"];
      }
    \end{dot2tex}
  \end{center}
  \caption{Semantic web content miner with addtional dereferencing of linked entities}
\end{sidewaysfigure}
